The landscape of chloroplast genome assembly tools

### 2019-06-06 bmc_article.tex
Update on Overleaf.

\section*{Methods}
\subsection*{Data availability}
Source code for all methods used is available at \url{https://github.com/chloroExtractorTeam/benchmark} and archived in zenodo under \zenodorepo{}.
All docker images are published on \url{https://cloud.docker.com/u/chloroextractorteam/} and are named with a leading \texttt{benchmark\_} (\cref{tab:dockerimages}).

@@

\subsection*{Our Setup}
We want to use a minimum of different parameter settings for all assembly programs to enable a fair comparison. 
Therefore, we decided to specify that all programs have to work based on two input files, representing a data set's forward (\texttt{forward.fq}) and reverse  (\texttt{reverse.fq}) sequence file in FASTQ format.
Depending on the assembler, output files with different names and locations are generated.
Those different files are copied and renamed to ensure that each assembly approach produces the same output file (\texttt{output.fa}). 
Additionally, we set an environment variable for all programs to control the number of allowed threads.
All three requirements (defined input file names, defined output file name, thread number control via environment variable) are ensured by a simple wrapper script (\texttt{wrapper.sh}).
Finally, for a maximum of reproducibility all programs have been bundled into individual docker images based on a central base image which provides all the required software. Those docker images were used for the recording of the consumption of computational resources.
Furthermore, all our docker images have been converted into singularity containers for the quantitative measurement on simulated and real data sets.
Singularity container were built from docker images for usage on a HPC-environment using Singularity v.2.5.2 \cite{kurtzer2017singularity}
All singularity containers were run on %the Julia-HPC 
\ak{I would note refer to Julia in the ms, maybe we could add it as a link}
Intel® Xeon® Gold \num{6140} Processors using a Slurm workload manager version 17.11.8 \cite{Jette02slurm}. Assemblies were run on \num{4} threads using \SI{10}{\gibi\byte} RAM with a time limit of \SI{48}{\hour}. 
\subsection*{Data}
\subsubsection*{Simulated}
To avoid suffering from sequencing errors and biological variances, we simulated perfect reads based on the \plant{At} (TAIR10) chloroplast assembly \cite{tair10}.
We used a sliding window approach with \texttt{seqkit} \cite{seqkit}. The exact commands are documented in \href{https://github.com/chloroExtractorTeam/benchmark/blob/master/03_representative_datasets.md}{\texttt{03\_representative\_datasets.md}} in \zenododataset{}.
For the final simulated data sets reads based on mixtures of the \plant{At} (TAIR10) core and chloroplast genome were generated with different ratios ( \num{0}:\num{1}, \num{1}:\num{10}, \num{1}:\num{100}, and \num{1}:\num{1000}).
Additionally, we generated data with different read lengths (\SIlist{150;250}{\basepair}). All data simulated contain exactly 2 million read pairs.

@@

\subsubsection*{Consistency}
\todo{lieber Consistency?}To ensure consistency of the obtained results, we randomly chose  \num{100} data sets, that in the previous runs resulted in outputs for most of the assemblers and run them again with the same parameters as before.
The resulting assemblies were scored again as described and the scores of the first and the second run were compared to each other
This information is important to assess the robustness of the different programs.

@@

\subsubsection*{Time requirements}
In terms of run time, massive differences between the different tools have been observed. Apart from tool-specific differences, input data and number of threads had huge impact. The observed run times varied from a few minutes to several hours (\cref{fig:performance_runtime}).
%\ak{did I miss Figure 1 ?, need to check figure numbering in the end}
Some assemblies failed to finish within the time-limit we set (\SI{48}{\hour}). 
On average the longest time to generate the assemblies was taken by \ioga{} and \fp{} followed by \oa{} and \go{} the most time efficient tool was \ce{}, which on average is a little faster than \np{} and \cassp{}.
%% multithreading

@@

\subsection*{Quantitative}
\subsubsection*{Simulated data}
The only assembler obtaining perfect results according to our score for the simulated data sets is \go{} (\cref{fig:simulated}).
\ioga{} and \cassp{} showed the worst performance, being unable to fully assemble a single chloroplast out of \num{14} runs.
\np{} performed second best with scores above \num{80} for all data sets, only failing to resolve the contigs into one single circular chloroplast assembly.
The overall performance \ak{of all assemblers ?} is best, when the input data consists purely of chloroplast reads.
Only \ioga{} and \cassp{} failed to deliver any results under this scenario once.
In general, no clear correlation between either length of the input reads or the ratio of core vs chloroplast reads and the performance of the different assemblers can be observed.

@@

\subsubsection*{Consistency}
Consistency was tested by re-running assemblies and comparison of the scores of two assemblies (\cref{fig:consistency}).
Replicates that did not produce an output were manually scored as \num{0}. 
\go{} was the only tool that succeeded in obtaining similar scores for all assemblies, without producing and completely unsuccessful assemblies for this subset of data.
Except for \fp{} all the other tools had at least one assembly that was unsuccessful in one run, but produced an output in the other.\ak{oben steht \go{} war als einziges immer gut, dann kommt except for \fp{} ?? does not seem to be consistent}
Notably \ioga{} appears to have a tendency to perform differently in independent runs.
Here, more than \SI{10}{\percent} of the assemblies failed in one run only.

Both \fp{} and \np{} tend to have minor changes in the assembly when the overall performance is comparably well, leading to the arrow-shaped scatter plots.\ak{link to figure scatter plots is missing}
\ce{} and \cassp{} appear to be the most robust assemblers, having only few deviations between the two runs.

@@

\subsection*{Guidelines for the end-user}
Given these results, our recommendation is to use \go{} as default option, and in case of failure \fp{} as backup solution.\ak{what if both lead to qualitative different results?, always trust \go{} , or could one compute a high-confidence assembly by combining different results ?, might be a point for future directions.}
If both programs fail, it is sensible to re-run \fp{} and additionally try \np{} and \oa{}.
This procedure maximizes the chance to effectively and efficiently recover the circular chloroplast genome from mixed genomic data.
If none of these four assemblers produce sensible results, a reference guided approach and tweaking of the default parameters, might be the solution.
Here, it is not possible to provide general guidelines, as the procedure will differ for different data sets.
For an automated approach, running \go{} and \fp{} in parallel appears to be a good trade-off between success rate and use of resources.
%\todo{Does the decision depend on some parameter of the data (read length, amount, taxonomic group, ...)?}

\subsection*{Ideas for future development}
%\ff{this section (lower priority)}
For further experiments, combining different components from different tools might be a promising approach.
For example, read scaling from \ce{} followed by an assembly by \go{} and finally the structural resolution with \fp{} could be a promising approach, combing the respective strength of the different tools.

@@

\section*{Competing interests}
The authors disclose that SP, NT, FF, and MJA are developers of \ce{}, one of the tools benchmarked in this article.
The authors exercised caution not to let this fact unfairly influence their judgment and recommendations.
JF, NT, and MJA are affiliated with the for-profit organization AnaLife Data Science.
\ak{and I think we should mention Analife as competing interest, normally people list all non-academic shareholder whatever stuff, just to be sure}\todo{Ich weiß nicht ob das so passt}

>>>>

\section*{Methods}
\subsection*{Data availability}
Source code for all methods used is available at \cite{github-benchmark-repo} and archived in zenodo under \zenodorepo{}.
All docker images are published on \cite{dockerhub-benchmark} and are named with a leading \texttt{benchmark\_} (\cref{tab:dockerimages}).

@@

\subsection*{Our Setup}
We want to use a minimum of different parameter settings for all assembly programs to enable a fair comparison. 
Therefore, we decided to specify that all programs have to work based on two input files, representing a data set's forward (\texttt{forward.fq}) and reverse  (\texttt{reverse.fq}) sequence file in FASTQ format.
Depending on the assembler, output files with different names and locations are generated.
Those different files are copied and renamed to ensure that each assembly approach produces the same output file (\texttt{output.fa}). 
Additionally, we set an environment variable for all programs to control the number of allowed threads.
All three requirements (defined input file names, defined output file name, thread number control via environment variable) are ensured by a simple wrapper script (\texttt{wrapper.sh}).
Finally, for a maximum of reproducibility all programs have been bundled into individual docker images based on a central base image which provides all the required software. Those docker images were used for the recording of the consumption of computational resources.
Furthermore, all our docker images have been converted into singularity containers for the quantitative measurement on simulated and real data sets.
Singularity container were built from docker images for usage on a HPC-environment using Singularity v.2.5.2 \cite{kurtzer2017singularity}
All singularity containers were run on %the Julia-HPC 
Intel® Xeon® Gold \num{6140} Processors using a Slurm workload manager version 17.11.8 \cite{Jette02slurm}. Assemblies were run on \num{4} threads using \SI{10}{\gibi\byte} RAM with a time limit of \SI{48}{\hour}. 
\subsection*{Data}
\subsubsection*{Simulated}
To avoid suffering from sequencing errors and biological variances, we simulated perfect reads based on the \plant{At} (TAIR10) chloroplast assembly \cite{tair10}.
We used a sliding window approach with \texttt{seqkit} \cite{seqkit}. The exact commands are documented in \href{https://github.com/chloroExtractorTeam/benchmark/blob/master/03_representative_datasets.md}{\texttt{03\_representative\_datasets.md}} in \zenododataset{}.
For the final simulated data sets reads based on mixtures of the \plant{At} (TAIR10) core and chloroplast genome were generated with different ratios ( \num{0}:\num{1}, \num{1}:\num{10}, \num{1}:\num{100}, and \num{1}:\num{1000}).
Additionally, we generated data with different read lengths (\SIlist{150;250}{\basepair}). All data simulated contain exactly 2 million read pairs.

@@

\subsubsection*{Consistency}
To ensure consistency of the obtained results, we randomly chose  \num{100} data sets, that in the previous runs resulted in outputs for most of the assemblers and run them again with the same parameters as before.
The resulting assemblies were scored again as described and the scores of the first and the second run were compared to each other
This information is important to assess the robustness of the different programs.

@@

\subsubsection*{Time requirements}
In terms of run time, massive differences between the different tools have been observed. Apart from tool-specific differences, input data and number of threads had huge impact. The observed run times varied from a few minutes to several hours (\cref{fig:performance_runtime}).

Some assemblies failed to finish within the time-limit we set (\SI{48}{\hour}). 
On average the longest time to generate the assemblies was taken by \ioga{} and \fp{} followed by \oa{} and \go{} the most time efficient tool was \ce{}, which on average is a little faster than \np{} and \cassp{}.
%% multithreading

@@

\subsection*{Quantitative}
\subsubsection*{Simulated data}
The only assembler obtaining perfect results according to our score for the simulated data sets is \go{} (\cref{fig:simulated}).
\ioga{} and \cassp{} showed the worst performance, being unable to fully assemble a single chloroplast out of \num{14} runs.
\np{} performed second best with scores above \num{80} for all data sets, only failing to resolve the contigs into one single circular chloroplast assembly.
The overall performance is best, when the input data consists purely of chloroplast reads.
Only \ioga{} and \cassp{} failed to deliver any results under this scenario once.
In general, no clear correlation between either length of the input reads or the ratio of core vs chloroplast reads and the performance of the different assemblers can be observed.

@@

\subsubsection*{Consistency}
Consistency was tested by re-running assemblies and comparison of the scores of two assemblies (\cref{fig:consistency}).
Replicates that did not produce an output were manually scored as \num{0}. 
\go{} was the only tool that succeeded in obtaining similar scores for all assemblies, without producing and completely unsuccessful assemblies for this subset of data.
Except for \fp{} all the other tools had at least one assembly that was unsuccessful in one run, but produced an output in the other.
%\ak{oben steht \go{} war als einziges immer gut, dann kommt except for \fp{} ?? does not seem to be consistent}
Notably \ioga{} appears to have a tendency to perform differently in independent runs.
Here, more than \SI{10}{\percent} of the assemblies failed in one run only.

Both \fp{} and \np{} tend to have minor changes in the assembly when the overall performance is comparably well, leading to the arrow-shaped scatter plots.
%\ak{link to figure scatter plots is missing}
\ce{} and \cassp{} appear to be the most robust assemblers, having only few deviations between the two runs.

@@

\subsection*{Guidelines for the end-user}
Given these results, our recommendation is to use \go{} as default option, and in case of failure \fp{} as backup solution.
%\ak{what if both lead to qualitative different results?, always trust \go{} , or could one compute a high-confidence assembly by combining different results ?, might be a point for future directions.}
If both programs fail, it is sensible to re-run \fp{} and additionally try \np{} and \oa{}.
This procedure maximizes the chance to effectively and efficiently recover the circular chloroplast genome from mixed genomic data.
If none of these four assemblers produce sensible results, a reference guided approach and tweaking of the default parameters, might be the solution.
Here, it is not possible to provide general guidelines, as the procedure will differ for different data sets.
For an automated approach, running \go{} and \fp{} in parallel appears to be a good trade-off between success rate and use of resources.

\subsection*{Ideas for future development}
For further experiments, combining different components from different tools might be a promising approach.
For example, read scaling from \ce{} followed by an assembly by \go{} and finally the structural resolution with \fp{} could be a promising approach, combing the respective strength of the different tools.

@@

\section*{Competing interests}
The authors disclose that SP, NT, FF, and MJA are developers of \ce{}, one of the tools benchmarked in this article.
The authors exercised caution not to let this fact unfairly influence their judgment and recommendations.
JF, NT, and MJA are affiliated with the for-profit organization AnaLife Data Science.
\ak{and I think we should mention Analife as competing interest, normally people list all non-academic shareholder whatever stuff, just to be sure}
\todo{Ich weiß nicht ob das so passt}
